---
title: "Convergent Cross Mapping"
output:
  html_notebook:
    df_print: paged
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

# Default colours for crossmapping to match those in slides.
xmap_colours <- c("X_Y" = "red",
                 "Y_X" = "blue")
```

```{r}
library(rEDM)
# library(nonlinearTseries)
library(tseriesChaos)
library(crqa)
library(tidyverse)
library(latex2exp)
source("R/coupled_logistic_map.R")
```

# Model example
A simple example of a nonlinear system with two variables is the coupled
logistic map. The two variables $X$ and $Y$ have internal dynamics depending on
the growth rates $r_X$ and $r_Y$. In addition there is interdependent
dynamics modeled as a causal effect of $X$ on $Y$ depending on
the coupling constant $\beta_{YX}$ and a coupling in the inverse direction depending on  $\beta_{XY}$.

The model is expressed by these two equations:
$$
  X_{n+1} = X_n ( r_X - r_X X_{n} - \beta_{XY}Y_n)
$$

$$
  Y_{n+1} = Y_n ( r_Y - r_Y Y_{n} - \beta_{YX}X_n)
$$

## Time series
The model is implemented in the function `coupled_logistic_map` which can generate time series of length `N` given initial values `x0`and `y0` of the two variables and values for all the parameters in the model. To avoid transient and possible ideosyncratic dynamics in the beginning, an optional parameter `N_skip` can be set that will skip the first `N_skip` data points and then run the model for an additional `N` generations.
```{r}
dimension <-  2
delay <- 1

# Note the parameters here are not the exact ones used in the slides
time_series <- coupled_logistic_map(
  x0 = 0.2,
  y0 = 0.6,
  rx = 3.65,
  ry = 3.8,
  bxy = 0,
  byx = 0.4,
  N = 1000,
  N_skip = 300
)
```

```{r plot time series}
ggplot(time_series, aes(x = time, y = X)) +
  geom_point(colour = "blue") +
  geom_line(colour = "blue") +
  xlim(c(1, 40)) +
  ylab("Population X") +
  theme_classic()

ggplot(time_series, aes(x = time, y = Y)) +
  geom_point(colour = "red") +
  geom_line(colour = "red") +
  xlim(c(1, 40)) +
  ylab("Population Y") +
  theme_classic()
```
## Attractor reconstruction

Reconstruct attractors from X and Y
```{r}
# Construct attractors using tseriesChaos::embedd()

MX <- embedd(X, m = dimension, d = delay)
MY <- embedd(Y, m = dimension, d = delay)

# Change the column names
colnames(MX) <- c("x", "y")
colnames(MY) <- c("x", "y")
```

```{r}
ggplot(as.data.frame(MX), aes(x = x, y = y)) +
  geom_point(color = "blue") +
  geom_point(data = as.data.frame(MY),
             color = "red") +
  theme_classic() +
  coord_fixed()
```


## Convergent cross mapping
```{r}
# Perform the cross mapping. This can take some time.
model_ccm <- CCM(dataFrame = time_series,
                 E = dimension, tau = -delay, Tp = 0,
                 columns = "Y", target = "X",
                 libSizes = "20 500 10", sample = 100,
                 random = TRUE,
                 showPlot = TRUE)


# Make sure the data frame has valid names (no colons)
colnames(model_ccm) <- make.names(colnames(model_ccm))
```

Plot the result ourselves
```{r}

model_ccm_long <- model_ccm %>% 
  rename(Y.MX = X.Y, X.MY = Y.X, L = LibSize) %>% 
  pivot_longer(cols = c("Y.MX", "X.MY"),
               names_to = "Crossmap",
               values_to = "Rho")

ggplot(model_ccm_long, aes(x = L, y = Rho, color = Crossmap)) +
  geom_line(size = 1) +
  # geom_point() +
  scale_color_manual(values = c("blue", "red"),
                     aesthetics = c("colour", "fill")) +
  ylab(TeX("$\\rho$")) +
  theme_classic()
```

# Sardine anchovy example
```{r}
 cmap <- CCM(dataFrame = sardine_anchovy_sst, E = 3, Tp = 0,
             columns = "anchovy", target = "np_sst",
             libSizes = "10 70 5", sample = 100,
             showPlot = TRUE)
```


# Breath, move, heart
```{r}
pair2 <- read_csv("Data/pair2/Study4_G2_T4_MovementCoop.csv")

# Down sample
pair2 <- pair2 %>%
 filter(row_number() %% 500 == 1)

# Subtract baseline
pair2$Resp1 <- pair2$Resp1 - mean(pair2$Resp1)
pair2$Resp2 <- pair2$Resp2 - mean(pair2$Resp2)

ggplot(pair2, aes(x = min, y = Resp1)) +
  geom_line() +
  geom_line(aes(x = min, y = Resp2), color = "red") +
  xlim(c(min(pair2$min), 10)) +
  theme_classic()

# Make z-scores
zscore <- function(vector) {
  (vector - mean(vector)) / sd(vector)
}

pair2$Resp1z <- zscore(pair2$Resp1)
pair2$HR1z <- zscore(pair2$HR1)

param <- list(method = "crqa", metric = "euclidean",
              maxlag = 20, radiusspan = 100, normalize = 0, rescale = 0,
              mindiagline = 2, minvertline = 2, tw = 0, whiteline = FALSE,
              recpt = FALSE , side = "both", datatype = "continuous",
              fnnpercent = 10, typeami = "mindip")

# This crashes on RStudio cloud
embed_param <- crqa::optimizeParam(pair2$Resp1z, pair2$HR1z,
                    par = param)

tseriesChaos::mutual(pair2$Resp1z, lag.max = 20, plot = TRUE)
tseriesChaos::mutual(pair2$HR1z, lag.max = 20, plot = TRUE)

tau <- 5

plot(false.nearest(series = pair2$Resp1z, 
                   m = 15, 
                   d = tau, 
                   t = 1, 
                   eps = 1))

plot(false.nearest(series = pair2$HR1z, 
                   m = 15, 
                   d = tau, 
                   t = 1, 
                   eps = 1))

E <- 4

ggplot(pair2, aes(x = min, y = Resp1z)) +
  geom_line() +
  geom_line(aes(x = min, y = HR1z), color = "red") +
  xlim(c(min(pair2$min), 10)) +
  theme_classic()


ggplot(pair2, aes(x = Resp1, y = Resp2)) +
  geom_point(size = 0.1, color = "grey", alpha = 0.4) +
  theme_classic()
```

```{r}
pair2_ccm_data <- pair2 %>% 
  select(Resp1z, HR1z)

# Get the row numbers and add them as the first column as required by CCM
idx <- as.numeric(rownames(pair2_ccm_data))
pair2_ccm_data <- cbind(time = idx, pair2_ccm_data)

ccm_out <- CCM(dataFrame = pair2_ccm_data, E = 4, tau = -3, Tp = 0,
               columns = "Resp1z", target = "HR1z",
               libSizes = "20 500 10", sample = 100,
               random = TRUE,
               showPlot = TRUE)

# Rename columns
ccm_out <- ccm_out %>%
  rename(L = LibSize,
         Resp_x_HR_1 = "Resp1z:HR1z",
         HR_x_Resp_1 = "HR1z:Resp1z")
```
Make our own plot
```{r}
ggplot(ccm_out, aes(x = L, y = Resp_x_HR_1)) +
  geom_line(colour = "red") +
  geom_line(aes(x = L, y = HR_x_Resp_1), colour = "blue") +
  ylab(TeX("$\\rho$")) +
  theme_classic()
```


```{r}
temperature <- read.table("Data/NASA/global_temp_index.txt",
                         skip = 5, header = FALSE)

colnames(temperature) <- c("Year", "T", "T_smooth")

co2 <- read.table("Data/NASA/co2_mm_mlo.txt",
                  comment.char = "#",
                  header = FALSE)


# Alternative
# https://datahub.io/core/global-temp#curl
# https://datahub.io/core/co2-fossil-global#data

```

